---
title: "Data Viz"
author: "Bernard A. Coles IV"
date: "2/14/2019"
output: html_document
---

clean_names(read_csv("data/prison_jail_1970_2016.csv")) the clean_names() changes things to lower case and underscores. It also helps to detect columns that contain numbers & spaces -> Good practice for uploading

gather() -> useful for making edgelists: (x = "key", y = "value", key = range of columns for series) Key(series, nodes) & value(counts, edges) are like creating two new columns -> think of it as narrowing the data table 

first step for working with new data is to try and reproduce the counts from the codebooks

vis_dat(<data.frame>) gives a cool map showing rough counts of missing data across variables and their class

gg_miss_var(<data.frame>) visualizes the missing data per variable
gg_miss_upset() visualized the missing data in a bar chart 
gg_miss_fct(x = data , fct = categorical variable) shows missing data across the categories of another variable  

miss_var_summary() missingness in columns -> variables
miss_case_summary() missingness in rows -> respondents 
*you can pipe these two functions at the end of a summary table for missingness 

Github repository -> The Quartz Guide to Bad Data : "Quartz/bad-data-guide"

naniar package contains replace_with_na() and useful vingettes on their website for handling missing data 

packrat - package that preserves the versions of packages you used to create your markdown 

sessionInfo() will tell you which packages are running as well as other plateform info so like you OS, etc. 

<function>_if() take action on a subset of variables such as class. This is called scoping if you need to google arguements (scope helpers)
example: summary_if(is.numeric(mydata))

matches() function might be useful for edgelist and node matching over aidr

select(world, everything()) will move the world column to the front. column shuffle 

filter_all(any_vars(is.na(.))) show me rows where there is missing data 

rowwise() can perform operations across rows which is useful creating new variables

every single map arguement in ggplot has a scale and they can conflict 

ggrepel is a good package for using text on plots 

gghightlight <- make certain features of the plot stand out from the rest 

facet_wrap( *reorder*(grouping), *value ordered on*)
facet_grid <- plots all of the groups overlapping on a single plot

```{r}
library(socviz)
library(tidyverse)
library(purrr)

data("organdata")

organdata %>%
  group_by(world) %>%
  summarize_if(is.numeric, mean, na.rm = T) %>%
  select(world, donors, pubhealth, roads) %>%
  select_all(tools::toTitleCase)

organdata %>%
  group_by(country) %>%
  summarize_if(is.numeric, funs(ave = mean, sd = sd), na.rm = T) %>% #funs(is similar to mutate but for creating colomns for means sds, etc)
  select(country, donors_avg, donors_sd, roads_avg, roads_sd) %>%
  arrange(desc(donors_avg))

out <- lm(donors ~ pop + gdp + roads, data = organdata)

#map is similar to the apply functions: iterates over all cases 
organdata %>%
  split(.$world) %>% #used split so that we didnt have to enclose all the maps in a mutate function 
  map(~lm(donors ~ pop + gdp + roads, data = .)) %>%
  map(summary) %>%
  map_dbl("r.squared")
```

###Class Notes - February 28, 2019
```{r}
library(socviz)
#setup_course_notes() all the code in the book by chapter - already downloaded
library(tidyverse)
library(devtools)
library(fs)
library(gapminder) #will be working with gapminder data
gapminder
```

###Class Notes - March 07, 2018
```{r}
a <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap)) +
  geom_line(color = "gray70", mapping = aes(group = country)) + 
  geom_smooth(size = 1.1, method = "loess", se = F) + 
  scale_y_log10(labels = scales::dollar) +
  facet_wrap(~ continent, ncol = 5)
a

b <- ggplot(data = gss_sm, mapping = aes(x = bigregion)) +
  geom_bar(mapping = aes(y = ..prop..))


```

###Class Notes - March 21, 2019 

#survey package 

#Maps: conceptually separating these steps into functional domains, easy to learn where to look for fixing a problem

```{r}
data(election)

election %>% 
  dplyr::select(state, total_vote, r_points, pct_trump, party, census) %>%
  sample_n(5)

setup_course_notes() #map.rmd <- notes from class in this file in the repository
```

```{r}
names <- nyc_license %>%
  group_by(borough) %>%
  tally() %>%
  mutate(freq = n / sum(n), pct = round(freq*100, 2)) %>%
  filter(animal_name != "Unknown")
```



































