---
title: "CQ Example"
author: "Kieran healy"
date: "1/24/2019"
output: html_document
---

# Setup

```{r onetime, eval = FALSE}

usethis::use_git()

```


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(lubridate)
library(socviz)
library(gapminder)

```

```{r local-functions}

## Calculate age in years, months, or days
calc_age <- function(birth_date, ref_date = Sys.Date(), unit = "year") {

    require(lubridate)

    period <- as.period(interval(birth_date, ref_date),
                        unit = unit)

    switch(unit,
           year = year(period),
           month = month(period),
           day = day(period),
           stop = "Unknown time unit. Choose year, month, or day.")

}

```

# CQ Data
## Load it
```{r load, include=FALSE}
filenames <- dir(path = "data/clean",
                 pattern = "*.csv",
                 full.names = TRUE)
filenames
data <- filenames %>% map_dfr(read_csv, .id = "congress")
data <- clean_names(data)
```

## Recodes
```{r recoding-1}
## The start and end and other date variables are CQ's and coded 01/03/1945 (i.e. m/d/y)
## We use lubridate's mdy() to convert them to date objects.
data <- data %>% 
  mutate(start = mdy(start), 
         end = mdy(end), 
         born = mdy(born), 
         death = mdy(death))
sessions <- as_tibble(data.frame(congress = 79:116,
                       start_year = seq(1945, 2019, by = 2),
                       end_year = seq(1947, 2021, by = 2)))

sessions <- sessions %>% 
  mutate(start_year = int_to_year(start_year, "01", "03"),
         end_year = int_to_year(end_year, "01", "03"))
data <- data %>% 
  mutate(congress = as.numeric(congress) + 78)
data
```


```{r jwervin}
data$born[data$last == "Ervin" & data$first == "Joseph"] <- ymd("1901-03-01")
```

Join sessions to main data.

```{r sessions}
data <- left_join(data, sessions)

## Character Vector of all name-dob combinations 
data <- data %>% 
  mutate(name_dob = paste(last, middle, first, born, sep = "--"))

## Uniques
name_dob <- as.character(unique(data$name_dob))

## create pid for unique name_dobs
persons <- tibble(pid = 1:length(name_dob),
                  name_dob = name_dob)

data <- left_join(data, persons)

data <- data %>% 
  mutate(start_age = calc_age(born, start_year))

data %>% 
  select(start_age)

```


William Joseph Green Jr and William Joseph Green III are father
and son; the latter succeeded to his father's seat. The raw data
have a mistaken entry for the son in the 79th (1945) congress,
giving his `start_age` as 6.

```{r wjgreen}

ind <- which(with(data, congress == 79 & ## 1945
                        last == "Green" &
                        middle == "Joseph" &
                        first == "William" &
                        suffix == "III")) ## but this is the son

data <- data[-ind,]


```

## New Variables

- `poc`: Person of color; binary
- `days_old`: N days old
- `months_old`: N months old
- `full_name`: Full Name
- `end_career`: Date of end career
- `entry_age`:  Age when first entered congress
- `yr_fac`: start_year as a factor rather than a date

```{r newvariables}
#data$poc <- recode(data$race, "White" = "White"", .default = "Non-White") like an if else statement gives white if yes, poc otherwise 

data <- data %>% 
  mutate(poc = recode(race, "White" = "White", .default = "Non-White"),
         days_old = calc_age(born, start_year, "day"), 
         months_old = calc_age(born, start_year, "month"),
         full_name = paste(first, last, suffix),
         entry_age = calc_age(born, start),
         yr_fac = factor(year(start_year)))
```

## poc: Person of color; binary
data$poc <- recode(data$race, "White" = "White", .default = "Non-White")

## days_old: N days old
data  <- data %>% mutate(days_old = calc_age(born, start_year, "day"))

## months_old: N months old
data  <- data %>% mutate(months_old = calc_age(born, start_year, "month"))

## full_name: Full Name
data  <- data %>% mutate(full_name = paste(first, last, suffix))

## end_career: Date of end career
data  <- data %>% mutate(end_career = mdy(end))

## entry_age:  Age when first entered congress
data <- data %>% mutate(entry_age = calc_age(born, start))

## yr_fac: start_year as a factor rather than a date
data$yr_fac <- factor(year(data$start_year))

```{r cleanwrite}

write_csv(data, "data/congress79_116.csv")

```


```{r}
mean_age <- data %>% filter(position == "U.S. Representative") %>%
    group_by(congress) %>%
    summarize(year = first(start_year), mean_age = mean(start_age, na.rm = TRUE))

data %>% filter(position == "U.S. Representative",
                    party %in% c("Democrat", "Republican")) %>%
    group_by(congress, sex, party) %>%
    summarize(year = first(start_year), mean_age = mean(start_age)) %>%
    filter(congress == 100)
```

```{r}
#Median age of representatives, by congress
median.age.rep <- data %>%
  filter(position == "U.S. Representative") %>%
  group_by(congress) %>%
  summarize(year = first(start_year), median.age = median(start_age, na.rm = T))

#Mean age of senators, by congress
mean.age.sen <- data %>% 
  filter(position == "U.S. Senator") %>%
  group_by(congress) %>%
  summarize(year = first(start_year), mean_age = mean(start_age, na.rm = T))

#Mean age of senators, by congress & party 
mean.age.scp <- data %>% 
  filter(position == "U.S. Senator") %>%
  group_by(congress, party) %>%
  summarize(year = first(start_year), mean_age = mean(start_age, na.rm = T))

#Age of oldest senator, by congress & party 
old.sen <- data %>% 
  filter(position == "U.S. Senator") %>%
  group_by(congress, party) %>%
  summarise(oldest = max(start_age, na.rm = T))
 
#Age of youngest representative, by congress & party 
young.rep <- data %>% 
  filter(position == "U.S. Representative") %>%
  group_by(congress, party) %>%
  summarise(youngest = min(start_age, na.rm = T))

#Count of N representatives for each congress, by party
count.rep <- data %>%
  filter(position == "U.S. Representative") %>%
  group_by(congress, party) %>%
  count()

#Frequency Table for Representatives, by Gender
gender <- data %>%
  filter(position == "U.S. Representative")
mytable <- table(gender$congress, gender$sex)
prop.table(mytable, 1)

#Class: January 31, 2018 starts here 

#How to do the table in Tidyr 
tidypct <- data %>%
  select(congress, sex) %>%
  group_by(congress, sex) %>%
  summarize(n = n()) %>%
  mutate(pct = (n/sum(n))*100)
tidypct

#top_n function will give you N maximum values from the variable 
#Example from class: gives you the top three job types before joining congress for each congress (originally start_year)
class.example <- data %>%
  select(congress, job_type1) %>%
  group_by(congress, job_type1) %>%
  summarise(n = n()) %>%
  mutate(pct = (n/sum(n)) * 100) %>%
  group_by(congress) %>%
  top_n(3, wt = pct) %>%
  arrange(desc(pct), .by_group = T) #arrange doesnt respect group_by have to tell it to 
```

attach(mydata)
mytable <- table(A,B) # A will be rows, B will be columns 
mytable # print table 

margin.table(mytable, 1) # A frequencies (summed over B) 
margin.table(mytable, 2) # B frequencies (summed over A)

prop.table(mytable) # cell percentages
prop.table(mytable, 1) # row percentages 
prop.table(mytable, 2) # column percentages

Using Gapminder & Producing Tables
```{r}
gapminder #just data doesnt come with functions? #library(readxl) provides functions for reading excel data 

edu #data from the socviz package #messy because years of schooling is split into like 5 varaibles 

#Using gather() and spread() to consolidate the data 
edu_t <- edu %>%
  gather(key = years.school, #creates new column years.school and puts the variables you want to gather in the cells
         value = freq, #will give counts for each category
         elem4:coll4) # what variables you want to gather
dim(edu_t) #2196 rows X 7 columns = long data 
dim(edu) #366 rows X 11 columns = compact data for display 
```

Table Joins
```{r}
#Conceptually its just merging tables 

#*left*_join() preserves values from the *left* table - things from the right with no matching columns will get NA's

#inner_join analogus to the intersection of the data so the only thing preserved is what they share not whats on the left or right

#full_join keeps everything from both - union that preserves everything

#anti_join & semi_join filtering joins

#Question 6 from Slack problems requires a join

senate <- data %>%
  filter(position == "U.S. Senator") %>%
  group_by(pid) %>% #the variable that uniquely identifies the congress people
  summarise(first = first(first),
            last = first(last),
            party = first(party),
            state = first(state),
            start = first(start),
            end = first(end))
house <- data %>%
  filter(position == "U.S. Representative") %>%
  group_by(pid) %>% #matching variable 
  summarise(first = first(first), 
            last = first(last), 
            start = first(start),
            end = first(end))

sen.house <- inner_join(senate, house, by = "pid") #matching variable used in by arguement here #anti_join would exclude people who have served in both the house and the senate
  
#stringr - str_detect(string, patter) - str_replace(string, pattern, replacement) these functions are pipeable
#string = variable - pattern = value - replacement = replace value 
#str_remove(variable, "value to delete") - useful for handling missing data? " NA$" kieran said use this in the "value to delete" 
#There is a link to exercises with regular expression on slack socviz 
#mac book patterns app -> test bed for regular expressions -> shows you what youre finding/replacing 
```











